A First Effort for a Distributed Segment-based Approach on Self-Assembled Nano Networks

### 2013-11-04 PAPER/implementation.tex
Synced to after-conferences updates

\section{Implementation Issues}
\label{sec:implementation}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.25\textwidth]{pictures/disr_imp.eps}
  \caption{\emph{DiSR} block architecture.}
 \label{fig:implementation}
\end{figure}

Let's so focus to the \emph{DiSR} block. In
Fig.~\ref{fig:implementation} is shown a possible sketch of this
block, which mainly consist in the following building elements:

\begin{itemize}
\item \emph{DBS block}: It Takes trace of the DBS, and how already
discussed in the Section~\ref{ssec:disr_dstruct}, there are six possible
configurations. These values can be codified using 3 bit implemented
with a 3-bit register and the required combinational logic. This block
is essentially designed as a state machine.
\item \emph{LED registers}: Is a set of registers useful for storing
the local environment data (LED) described in
Section~\ref{ssec:disr_dstruct}. In particular in
fig.~\ref{fig:implementation}, are highlighted two tables representing
the \emph{link\_visited[]} and \emph{link\_tvisited[]} values. This
block consists of an array of registers (one for each
direction) that can be implemented with an SRAM, in which are presents
two specific fields  for representing the \emph{segID}. 
\item \emph{Control circuitry}: This circuitry reads data
from the fields of the incoming packet, LED registers and the
\emph{DBS}, updating \emph{LED} and
\emph{DBS} data when required. Then, the resulting \emph{Ctrl Out} output  drives the other
communication resources (such as node transceivers) for actuating the
DiSR routing operation.
\end{itemize}

As discussed above, one of the main design challenge of DNA
Self-Assembled systems is the limited resource available for
implementing both computations and routing decisions for each node.
For this reason, assuming a budget of $10^4$ CNFETs for each network's
node~\cite{liu_jetcs} we estimated the required resource for
implementing the entire DiSR block. A not optimized behavioural
HDL of the DiSR circuitry has been written and synthesized at gate-level.
Considering the specific layout of each single logic elements ( NAND,
full-adder, latch etc.), it has been possible to get a rough estimate
of the number of transistors necessary for implementing the DiSR
logic.  Figure~\ref{fig:imp_trend} shows the results of synthesis in
terms of number of devices (CNFETs) versus the number of network nodes
while the network scales up from $10\times10$ to $100\times100$ nodes.
\begin{figure}
  \centering
  \includegraphics[width=0.25\textwidth]{pictures/imp_rslt.eps}
  \caption{DiSR synthesis results: N is the number of network's node.}
 \label{fig:imp_trend}
\end{figure}
In particular, observing the Fig.\ref{fig:imp_trend} can 
be pointed out that the proposed implementation occupy  about 20\% of the node budget. 
Further, more than the absolute number of devices itself, it is interesting to
observe that, while the node have a squared increment, the circuitry
complexity which implements the DiSR algorithm increases with a slowly
growing trend.  
An intuitive explanation of this is the relatively simple logic of
DiSR, whose behaviour is almost code on scalable storage structures.
For example, the number of registers implementing the
\emph{link\_visited[]} and \emph{link\_tvisited[]} table follow the
logarithmic function:
\begin{equation}
  \label{eq:imp_trend}
  N_{reg}=2\cdot N_{port} \cdot log_2(N)
\end{equation}
where $N_{port}$ is the number of the router's ports, $N$ is the number 
of network's nodes while the term $2$ takes in account that \emph{link\_visited[]} and 
\emph{link\_tvisited[]} are two separate set of registers. While this number is fixed by 
the number of network's node, the control circuitry could be optimized applying more effort 
in future designs of this building block.

>>>>

\section{Hardware Implementation}
\label{sec:implementation}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/implementation.eps}
  \caption{\emph{DiSR} block architecture.}
 \label{fig:implementation}
\end{figure}

To give a basic estimate of the overhead needed we will focus on
DiSR-specific components, that is, the control logic and configuration
registers needed to implement DiSR. In
Figure~\ref{fig:implementation} is shown a sketch of this block, which
mainly consist in the following building element: 
\begin{itemize}
	\item \textbf{DBS block:} It takes trace of the DBS state machine, 
	consisting of a 3-bit register (to cover six DBS values) and the 
	required combinational logic. This block receives signals from 
	stored packets to decode the packet type, and from 
	the control circuitry to change its status.
	The bootstrap node is selected by setting the boot signal to high
	during the initialization phase.  
	\item \textbf{LED registers:} A set of registers storing 
	the Local Environment Data (LED) listed in the section
	\ref{sec:disr_concepts}. It should be pointed out that, while the \emph{visited} 
	and \emph{tvisited} data must be implemented with two distinct
	elements, the \emph{tvisited} and \emph{visited} table can be implemented with 
	a single RAM file for storing the segment id of each link.
	This can be done adding a special register named \emph{Tflag}
	indicating whether the \emph{SegID} value refers to \emph{visited} or to 
	\emph{tvisited}. When SegID's register is set to 0, it means that the link
	is neither visited nor tvisited.

    \item \textbf{Control circuitry:} This circuitry reads data from
	the incoming packet, the LED registers and the DBS,
	updating them when required. This block drives the LED register 
	for storing or, for reading the LED data according with
	\emph{write enable (WE)} or \emph{read enable (RE)}
	signals. Then, the resulting Ctrl-Out output drives the other
	communication resources for actuating the DiSR routing operation.
\end{itemize}

Each building block reacts with the rising edge  of
the\emph{clock signal} shown in Figure. An asynchronous system
\emph{reset} signal is also present for restoring all registers to
their default values. When this last signal is
set, the DBS state machines for each network's node are set as
\emph{Free}.

As discussed above, one of the main design challenge of DNA
Self-Assembled systems is the limited resource available for
implementing both computations and routing decisions for each node.
Assuming a budget of $10^4$ CNFETs for each network's
node~\cite{liu_jetcs}  we estimated the required resource for
implementing the entire DiSR block.  A not optimized RTL description
(with an hardware description language)  of the DiSR circuitry has
been written and synthesized at gate-level. Considering the specific
layout of each single logic elements (NAND, full-adder, latch etc.),
it has been possible to get a rough estimate of the number of
transistors necessary for implementing the DiSR logic.

Figure~\ref{fig:imp_trend} shows the results of synthesis in
terms of number of devices (CNFETs) versus the number of network nodes
while the network scales up from $10\times10$ to $100\times100$ nodes.
\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/imp_rslt.eps}
  \caption{Number of network's node (N) vs number of CNFETs.}
 \label{fig:imp_trend}
\end{figure}
In particular, observing the Figure~\ref{fig:imp_trend} can 
be pointed out that the proposed implementation occupy  about 17\% of the node budget. 
Further, more than the absolute number of devices itself, it is interesting to
observe that, while the node have a squared increment, the circuitry
complexity which implements the DiSR algorithm increases with a slowly
growing trend.  
An intuitive explanation of this is the relatively simple logic of
DiSR, whose behaviour is al4ost code on scalable storage structures.
For example, the number of registers
implementing the \emph{link\_visited[]} and \emph{link\_tvisited[]}
table follow the logarithmic function $N_{reg}=N_{port} \cdot
log_2(N)$ where \emph{Nport} is the number of the router’s ports, N is
the number of network’s nodes.  While $N_{reg}$ is fixed by the number
of network’s node,  the control circuitry could be optimized applying
more effort in future designs of this building block.

### 2013-11-04 PAPER/introduction.tex
Synced to after-conferences updates

% TODO: re-enable on long paper
%Exploring long term alternatives to the CMOS technology is gaining
%more and more revelance as the scaling trend of such devices keeps
%introducing new challenges. Power density, defect tolerance, testing
%costs and wire delays are only a few of the many critical aspects
%involved~\cite{itrs12}. While software parallelism and multicore
%approaches~\cite{horowitz2004, mudge2001, powell2009} are partially mitigating the impact of such
%constraints on performances, it is likely that the growing computing demand will
%eventually need even more radical architectural modifications and new
%paradigms in order to address the Computer Design challenges of the next
%decades.

In the last years, Self-assembled nanoscale architectures~\cite{winfree1998, yan2003}
are emerging as a promising technology due their tera/peta scale of
integration, defect tolerance and huge potential computing
capabilities. These technologies are certainly still at
their early stage of development, however different laboratory demos and
prof-of-concepts architectures have been presented~\cite{patwardhan2004, patwardhan2006_1, pistol2009}.
%, patwardhan2006
The main idea behind this approach is to exploit the physical regularity and
stability of DNA structures in order to create a scaffold onto which
nano-devices (e.g. nanowires and CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be
attached. This can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold.
%(see figure~\ref{fig:dna_tag}.
%
%\begin{figure}
%  \centering
%    \includegraphics[width=0.30\textwidth]{pictures/dna2b.eps}
%  \caption{DNA tag matching}
%  \label{fig:dna_tag}
%\end{figure}
A detailed description of the chemical properties involved is far
beyond the scope of this paper (see also~\cite{braun1998, seeman1999}), so we will focus on
the three main challenges that this new fabrication process introduces
in Computer Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.

@@

These aspects of a DNA-self assembly process lead to some important
implications to be addressed when approaching to the design of
computer architecture: a computational model should be founded on a
distributed network of small computing and storage nodes, randomly
placed and interconnected (see figure~\ref{fig:nana}).
Proof-of-concept of such kind of computational model can be found for
example in~\cite{patwardhan2006_1}.  Since no regularity can be
assumed in such network, a topology agnostic strategy that avoids
deadlock should be adopted in order to route packets (e.g.
instructions and data operands) among nodes.  It is important to
underline that, altougth we can assume an architecture similar to the
one depicted in Figure~\ref{fig:nana} as background, the topology discovery
problem faced in this work is more general and not strictly dependent on any of the
underlying computing architecture.

%where instructions and data operands travel
%on packets and routed using a topology-agnostic strategy that avoids
%deadlock, 
\begin{figure}
  \centering
    \includegraphics[width=0.20\textwidth]{pictures/dna1.eps}
  \caption{Irregular topology of computing and storage elements}
  \label{fig:nana}
\end{figure}

@@

\section{Related Works}
\label{sec:related_works}
Several works~\cite{sancho2002, skeie2002, skeie2004, koibuchi2003} have been presented addressing the problem of topology
agnostic routing. While they show interesting
performances, the requirement for virtual channels resources 
to avoid deadlock makes them not suitable for the specific DNA self
assembled networks we are assuming as background.
Other topology independent algorithms~\cite{schroeder1991, koibuchi2001, cherkasova1996} do not use virtual channels to
achieve deadlock freedom, but turn prohibitions. 
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploits the creation of a spanning tree of the
topology, then placing bidirectional restrictions by avoiding a packet
to traverse the same link in both up and down directions.
While the hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed,
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{dally1994, duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, sui2000, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010, kim2006, park2006}.

>>>>

 TODO: re-enable on long paper
Exploring long term alternatives to the CMOS technology is gaining
more and more revelance as the scaling trend of such devices keeps
introducing new challenges. Power density, defect tolerance, testing
costs and wire delays are only a few of the many critical aspects
involved~\cite{itrs12}. While software parallelism and multicore
approaches~\cite{horowitz2004, mudge2001, powell2009} are partially mitigating the impact of such
constraints on performances, it is likely that the growing computing demand will
eventually need even more radical architectural modifications and new
paradigms in order to address the Computer Design challenges of the next
decades.

In the last years, Self-assembled nanoscale architectures~\cite{winfree1998, yan2003}
are emerging as a promising technology due their tera/peta scale of
integration, defect tolerance and huge potential computing
capabilities. These technologies are certainly still at
their early stage of development, however different laboratory demos and
prof-of-concepts architectures have been presented~\cite{patwardhan2004, patwardhan2006_1, pistol2009}.
%, patwardhan2006
The main idea behind this approach is to exploit the physical regularity and
stability of DNA structures in order to create a scaffold onto which
nano-devices (e.g. nanowires and CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be
attached. This can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold.
A detailed description of the chemical properties involved is far
beyond the scope of this paper (see also~\cite{braun1998, seeman1999}), so we will focus on
the three main challenges that this new fabrication process introduces
in Computer Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.

@@

\begin{figure}
\centering
\begin{tabular}{cc}
    \includegraphics[width=0.40\textwidth]{pictures/dna2b.eps} &
    \includegraphics[width=0.40\textwidth]{pictures/dna1_complex2.eps} \\
 (a) & (b)
 \end{tabular}
  \caption{DNA tag matching and Resulting Irregualr Topology}
  \label{fig:nana}
\end{figure}
These aspects of a DNA-self assembly process lead to some important
implications to be addressed when approaching to the design of
computer architecture: a computational model should be founded on a
distributed network of small computing and storage nodes, randomly
placed and interconnected (see figure~\ref{fig:nana}).
Proof-of-concept of such kind of computational model can be found for
example in~\cite{patwardhan2006_1}.  Since no regularity can be
assumed in such network, a topology agnostic strategy that avoids
deadlock should be adopted in order to route packets (e.g.
instructions and data operands) among nodes.  It is important to
underline that, altougth we can assume an architecture similar to the
one depicted in Figure~\ref{fig:nana} as background, the topology discovery
problem faced in this work is more general and not strictly dependent on any of the
underlying computing architecture.

@@

\section{Related Works}
\label{sec:related_works}
Several works~\cite{sancho2002, skeie2002, skeie2004, koibuchi2003} have been presented addressing the problem of topology
agnostic routing. While they show interesting
performances, the requirement for virtual channels resources 
to avoid deadlock makes them not suitable for the specific DNA self
assembled networks we are assuming as background.
Other topology independent algorithms~\cite{schroeder1991, koibuchi2001, cherkasova1996} do not use virtual channels to
achieve deadlock freedom, but turn prohibitions. 
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploits the creation of a spanning tree of the
topology, then placing bidirectional restrictions by avoiding a packet
to traverse the same link in both up and down directions.
While the hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed,
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{dally1994, duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, sui2000, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010, kim2006, park2006, ebrahimi2013}.

### 2013-11-04 PAPER/simulation.tex
NocArc review fixes
Synced to after-conferences updates

% TODO: long paper
%\subsection{Other optimisations}
%Some optimisation parameters, which demostrated to improve the DiSR
%results have been fixed to some good values, but are not subject to
%further investigation; once again the focus here is not the
%optimal set of segments, but just demostrating a working approach. 
%These are optimisations are:
%\begin{itemize}
%\item{cycle\_links}: max number of retries across the set of links of
%each node. While searching for a free link due an incoming request,
%the same request is cancelled after a given number of tries. This
%gives to preceding node change to test a different paths. Default is set to 1.
%\item{bootstrap\_timeout}: number of time units that a bootstrap node
%should wait before assuming that a livelock in the starting segment
%process has occurred. In the worst case, we can imagine that longest
%path required is the one returning to the bootstrap node after having
%travelled across all the links. So, althought this is just an extreme
%situation, a good upper limit can be safely be set to NxN.
%\item{bootstrap\_immunity}: in order to avoid the failure of the whole DiSR
%setup process, a bootstrap node should not have defective links.
%Enabling this optimisation, a bootstrap node is immune to defects.
%We may think of a pre-bootstrap phase that properly selects (from upper
%layer via) a bootstrap node which is tested as properly connected. We
%enabled this optimisation, however empirical tests showed us that only
%simulations using bootstrap nodes placed on edges would be heavily
%affected, since these are nodes starting with a lower number of links,
%e.g. corner nodes could only have two connected direction, so even a
%defective link could prevent starting segment packet to return to
%bootstrap to close the loop a create the segment.
%\end{itemize}

@@

Figure~\ref{fig:results} collects the results in terms of node
coverage and latency at different network sizes, defect rates and
bootstrap injection points.  In particular graphs (a) and (b) show
node coverage for DiSR and RPF tree based approach respectively. While
the first aim of DiSR is not to reach the optimal coverage, we still
can observe a quite good performance as compared to tree based
approach. Note that defect rates beyond 25\% lead to many disconnected
regions of nodes that this first of DiSR cannot handle; however, these
level of defects should be considered as worst case scenarios, so the
achieved coverage of $0.5$ is satisfying results for this first
version of DiSR. As regards the the network size it seems to have a
limited impact when defect rate do not introduce to much disconnected
regions.

The number of cycles required to complete the segment mapping
process is shown in latency graph (c). Comparison against tree-based
has been omitted for space reasons, however similar numbers (in terms
of upper bound) have been reported
(see~\cite{Patwardhan05evaluatingthe} for details). What it's more
interesting is to observe is how DiSR latency scales with network
size. For example, going from 900 to 2500 nodes, at the medium defect
rate of $0.15$, leads to an increase from 3000 to 4500 cycles. Note how
the effect of defect rate is increasing until the threshold of $0.25$
is reached, meaning that while DiSR finds it more difficult to
complete the process due increasing defective paths, it still discover
new segments when let running for a more extended amount of cycles.
After that threshold, the impact of entire disconnected regions
becomes predominant and DiSR more quickly complete the process
covering the nodes that can be reached.

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.22\textwidth]{pictures/set1.eps} & 
\includegraphics[width=0.22\textwidth]{pictures/coverage.eps} \\
(a) & (b) \\
\includegraphics[width=0.22\textwidth]{pictures/set2.eps} & 
\includegraphics[width=0.22\textwidth]{pictures/set3.eps} \\
(c) & (d)
\end{tabular}
\caption{DiSR (a) and RPF (b) node coverage (c) Latency and effect
of bootstrap node (d)}
\label{fig:results}
\end{figure}

Finally, picture (d) visually represent  the stability of the approach
against different choices of the bootstrap node in a 10x10 network.
This is an important aspect to evaluate when considering that one of
the main advantages of DiSR against all the tree based approaches
should be the ability of choosing whatever bootstrap node, without
having to care of any particular \emph{root-node role} assumed in the future.
In other words, after segments have been established, the bootstrap node
is like every other node, i.e. it is not center of a structure, and it
is not an hotspot for the traffic distribution. The results in terms
of coverage,  shown for low, medium and high defect profiles, seems to
demonstrate a relatively limited impact of the bootstrap choice in
the low/medium scenarios, while a 30\%instability is found for very high defect
rates. This also sounds acceptable, since when lot of defective nodes
are present, the particular position of the bootstrap choice could
lead to completely different evolution in the DiSR setup process.

>>>>

% TODO: long paper
\subsection{Other Optimisations}
Some optimisation parameters, which demostrated to improve the DiSR
results have been fixed to some good values, but are not subject to
further investigation; once again the focus here is not the
optimal set of segments, but just demostrating a working approach. 
These are optimisations are:
\begin{itemize}
\item{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due an incoming request,
the same request is cancelled after a given number of tries. This
gives to preceding node change to test a different paths. Default is set to 1.
\item{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that longest
path required is the one returning to the bootstrap node after having
travelled across all the links. So, althought this is just an extreme
situation, a good upper limit can be safely be set to NxN.
\item{bootstrap\_immunity}: in order to avoid the failure of the whole DiSR
setup process, a bootstrap node should not have defective links.
Enabling this optimisation, a bootstrap node is immune to defects.
We may think of a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimisation, however empirical tests showed us that only
simulations using bootstrap nodes placed on edges would be heavily
affected, since these are nodes starting with a lower number of links,
e.g. corner nodes could only have two connected direction, so even a
defective link could prevent starting segment packet to return to
bootstrap to close the loop a create the segment.
\end{itemize}

@@

In this section we analyze the results in terms of node
coverage and latency at different network sizes, defect rates and
bootstrap injection points.  In particular
Figure~\ref{fig:results_coverage} shows
node coverage for DiSR and RPF tree based approach respectively. While
the first aim of DiSR is not to reach the optimal coverage, we still
can observe a quite good performance as compared to the tree based
approach. Note that defect rates beyond 25\% lead to many disconnected
regions of nodes that DiSR currently cannot handle.
For example Figure~\ref{fig:net} show a 30x30 network with a 25\%
defect in which the bottom-left part cannot be reached due
disconnected regions. Note that remaining defective nodes belonging to
connected regions are successfully surrounded by DiSR coverage; in any
case, these defect levels should be considered as worst case scenarios, so the
achieved coverage of $0.5$ is a satisfying result for this first
version of DiSR. On the other hand, the network size seems to have a
limited impact when defect rate do not introduce too much disconnected
regions. 
\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{pictures/set1.eps} & 
\includegraphics[width=0.48\textwidth]{pictures/coverage.eps} \\
(a) & (b) 
\end{tabular}
\caption{DiSR (a) and RPF (b) node coverage}
\label{fig:results_coverage}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.70\textwidth]{pictures/net.ps}
\caption{Covered regions in a 30x30 network with 25\% defects}
\label{fig:net}
\end{figure}

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{pictures/set2.eps} & 
\includegraphics[width=0.48\textwidth]{pictures/set2_rpf.eps} \\
(a) & (b)
\end{tabular}
\caption{Latency of DiSR (a) vs tree based RPF (b) }
\label{fig:results_latency}
\end{figure}

The number of cycles required to complete segment mapping process
is shown in Figure~\ref{fig:results_latency}. In this case comparison
against tree-based shows better (lower) values at different defect
rates.  Rather than the absolute numbers, what it's more interesting
to observe is how DiSR latency scales with network size. For example,
going from 900 to 2500 nodes, at the medium defect rate of $0.15$,
leads to an increase from 3000 to 4500 cycles. It should be noticed
also how the effect of defect rate is increasing until the threshold
of $0.25$ is reached, meaning that until that limit DiSR finds it more
and more difficult to complete the process due increasing defective
paths, but still discover new segments when let running for a more
extended amount of cycles. This behavior is not reported in the RPF
based approach, meaning that a tree based approach, although starting
from higher values, is less affected by defect rates (when low rates
are considered).
After the $0.25$ threshold, the impact of entire disconnected regions
becomes predominant and both approaches become faster in completing
the covering process, since far less nodes can be actually reached.

Finally, Figure~\ref{fig:results_bootstrap} visually represents  the
stability of the approach against choices of a different bootstrap
node in a 10x10 network.  This is an important aspect to evaluate when
considering that one of the main advantages of DiSR against all the
tree based approaches should be the ability of choosing whatever
bootstrap node, without having to care about the role assumed in the
future by the chosen root node.  In other words, after segments have
been established, the bootstrap node is like every other node, i.e. it
is not center of a structure, and it is not an hotspot for the traffic
distribution. The results in terms of coverage,  shown for low, medium
and high defect profiles, demonstrate a relatively limited
impact of the bootstrap choice in the low/medium scenarios, while a
$30\%$ instability is found for very high defect rates. This also
sounds acceptable, since when a lot of defective nodes are present, the
particular position of the bootstrap node could lead to a completely
different evolution in the DiSR setup process.

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth]{pictures/set3.eps}
\caption{Effect of bootstrap node}
\label{fig:results_bootstrap}
\end{figure}
